---
title: "Recurrent Pattern of Staph Infection among Chindren in GA from 2002-2019"
author: "Shifan Yan"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: scroll
---

```{r setup, include=FALSE}
pacman::p_load(
  tidyverse, table1, cowplot, plotly, htmlwidgets, stringr, DT, flexdashboard
)

staph_p <- read.csv("person-level_2002-2019.csv")
staph_inf <- read.csv("infection-level_2002-2019.csv")

staph_p <- staph_p %>%
  left_join(
    staph_inf %>%
      group_by(MRN) %>%
      summarise(
        inf_count = n(),
        mssa_count = sum(Staph_Type == "MSSA"),
        mrsa_count = sum(Staph_Type == "MRSA")
      ),
    by = join_by(MRN)
  ) %>%
  filter(inf_count > 1)
```

Data Visualization
=====================================

Column {data-width=500, .tabset}
-----------------------------------------------------------------------

### Distribution of Infection Counts for Recur Patients in 2002-2009

```{r}
p <- ggplot(staph_p, aes(x = factor(inf_count), fill = Staph_Type_Status)) +
  geom_bar(position = "dodge") +  # Create side-by-side bars grouped by Staph_Type_Status
  labs(
    x = "Infection Count",
    y = "Number of Patients",  
    fill = "Staph Type Status"
  ) +
  theme_cowplot() +
  theme(
    axis.text.y = element_text(size = 9),
    axis.text.x = element_text(size = 9, angle = 45, hjust = 1),               # Tilt x-axis labels for readability
    legend.text = element_text(size = 9),                            # Make legend text smaller
    legend.title = element_text(size = 10),                          # Make legend title slightly larger
    legend.key.size = unit(0.2, "cm")                                # Shrink legend key size
  )


# turn to interactive
ggplotly(p, tooltip = "text") %>%
  layout(
    legend = list(orientation = "h", x = 0, y = -0.2),  # move legend to bottom
    title = list(x = 0)  # left-align title
  )
#the tooltip = "text" code ensures a box of text is shown when hovering
#over a datapoint
```

### Key Takeaways

 - Most recurrent patients had **2 episodes**. 
 - **MRSA is more common** than MSSA among children with repeated infections. 
 - The interactive plot allows exploration by infection count and type status.


Column {data-width=500 .tabset}
-----------------------------------------------------------------------

### Demographic Pattern of Recurrent Staph Infection of 2002-2019

```{r}
staph_p <- staph_p %>% 
  mutate(
    Recurrent_Status = factor(Recurrent_Status, levels = c("Recurrent after 365 days", "Recurrent within 365 days", "Recurrent after 365 days and within 365 days", "Single Episode")),
    Race = str_to_title(Race) %>% factor(levels = c("White", "Black", "Other")),
    Ethnicity = str_to_title(Ethnicity) %>% recode("Other" = "Other / Unknown") %>% factor(levels = c("Hispanic", "Non-Hispanic", "Other / Unknown")),
    Insurance_Status = str_to_title(Insurance_Status) %>% recode("Other" = "Other / Unknown") %>% factor(levels = c("Public", "Private", "Self Pay", "Other / Unknown")),
    Staph_Type_Status = factor(Staph_Type_Status, levels = c("MRSA", "MSSA", "Shifting"))
  )

summary_table <- table1(
  ~ Sex + Age + Race + Ethnicity + Insurance_Status + Staph_Type_Status | Recurrent_Status,
  data = staph_p, 
  overall = "Total")

# table1 to data.frame
summary_df <- as.data.frame(summary_table)

# interactive form
datatable(summary_df, 
          class = 'compact',
          extensions = 'Buttons', 
          options = list(
            dom = 'Bfrtip',  # export button
            buttons = c('copy', 'csv', 'excel', 'pdf', 'print'), 
            pageLength = 50,  
            scrollX = TRUE  
          ),
          rownames = FALSE)  # hide row number

```

### Key Takeaways

 - **Older age at first recurrence** seen in patients with **later recurrence**(p < 0.001)
 - **Race & Ethnicity differences** (p < 0.001): Recurrent infections more common among **White** and **non-Hispanic** children
 - **Insurance status** associated with recurrence (p < 0.001): Recurrent cases had **more public insurance**, fewer **self-pay patients**
 - **~25%** of all recurrent cases showed **Staph type shift** (MRSA ↔ MSSA) 


Description
===================================== 

#### **Project Motivation and Sample Description**

This project uses a confidential dataset of Staphylococcus infections from electronic health records between 2002 and 2019, collected from over 20,000 patients at Children’s Healthcare of Atlanta. It focuses on **recurrent Staph infections** across different demographic groups.

After filtering for community-onset infections and excluding patients younger than 19 years old who live in Georgia, the study population consists of **20,361 patients**.

The dataset includes:
- **Infection Count**
- **Demographic Variables**: Age, Sex, Race, Insurance Status
- **Staph Type**: MRSA, MSSA, and Shifting types

#### **Research Questions**

1. How does recurrence status vary across demographic groups?
2. Are certain groups more prone to recurrence?
3. How do Staph types correlate with recurrence?

#### **Public Health Relevance**

This dashboard supports healthcare providers by helping identify **high-risk populations**, ultimately guiding interventions and policies targeting **recurrent Staph infections in pediatric populations**.

#### **GitHub**

This dashboard has been display in my [GitHub](https://github.com/inkestonine/Data555_Dashboard)